Four reasons your build is slow

### 2019-01-17 paper.tex

Some projects may exhibit ample potential parallelism but may not achieve that potential
parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
These calculations where made from a build of LLVM that was run without strace.

A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
LLVM should build in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
built with ``-j 32'' the observed speedup was slightly over 18; see chart~\ref{fig:llvm}.  Using work and
span our tool can predict the potential speedup a build will experience using Brent's theorem
\cite{brents}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
P processors, \begin{math} T_P \leq T \end{math}

>>>>

Some projects may exhibit ample potential parallelism but not achieve that potential
parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours, and
a predicted span of 305 seconds, or 5 minutes, measured from a build of LLVM that was run without
strace. With a work of 3.5 hours and a span of 5 minutes, LLVM has a parallelism of about 41.

A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
LLVM would build in about 5 minutes.  Our observations however, did not match this expectation.
In chart \ref{fig:llvm} we see two different observed speedups, indicted by the squares and
triangles.  The squares indicate the observed speedup of LLVM when run on a machine with 32 cores,
with \emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the observed speedup when run
on a different machine, with 32 real cores and 32 hyperthreads, with \emph{make -j n} from where
n is most of the numbers between 1 and 32.  The dashed line indicates the upper predicted speedup,
which is the speedup the project can't exceed.  It is apparent from the chart that LLVM fails to
achieve this maximum predicted speedup.  The solid line represents the predicted lower speedup,
which is the speedup the project should achieve at minimum.  And it appears in the chart that
LLVM narrowly achieves this speedup, when built with \emph{make -j n} for most n.

The lower and upper speedup bounds are calculated using Brent's theorem \cite{brents}, which
specifies upper and lower bounds on the time, $T_P$ it will take a parallel algorithm to run on P
processors.

\begin{equation*}
  \frac{work}{P} \geq T_P \leq \frac{work}{P} + span 
\end{equation*}
